---
title: "Model of Perception and Action"
author: "Kiri Koppelgaard"
date: "June 3, 2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages}
###Loading packages
library(pacman)
p_load(ggplot2, tidyverse, signal)

```

#Load the data set and rename the columns to sample, time, x, y, z

```{r loading data}
###Loading data
data <- read.table("Cognitive Science/4. Semester/Models for perception and action/Exam/Models_of_action_perception/timsigF_211_1_1_1_short.txt", sep = ",", header = FALSE)

#Removing row names
data <- data[2:5]

```

```{r renaming columns}
names(data) <- c("time", "x", "y", "z") #Renaming columns
```

#Plot x versus z and add meaningful labels
```{r}
ggplot(data, aes(x, z))+
  geom_path()+ #adding geom to track path
  theme_bw()+ #Changing background colour
  labs(y="Movement on z-axis", x = "Movement in x-axis") #Adding labels
```


#Correct the z dimension by flipping all values so that more positive ones are shown upwards, i.e. like the movement was actually performed. Make a new plot of x and z
```{r}
#Flipping the z-axis by subtracting the values from 0, this reverts the signs
data$flipped_z <- 0-data$z

ggplot(data, aes(x, flipped_z))+
  geom_path()+ #adding geom to track path
  theme_bw()+ #Changing background colour
  labs(y="Movement on z-axis", x = "Movement in x-axis") #Adding labels
```

#Calculate z velocity and plot it against time
Velocity can be calculated using the following formula: Velocity = distance difference / time difference. Thus, we need to calculate the distance difference and time difference, before calculating velocity. 

```{r calculating z velocity}
#Calculating distance difference on the z-axis
data$distance_dif_z <- data$z-lag(data$z) #subtracting a given row with the value of the previous row, the lag function supplies the previous row  

#Calculating time difference
data$time_dif_z <- data$time-lag(data$time) #subtracting a given row with the value of the previous row

#Calculating velocity
data$velocity_z <- data$distance_dif_z /data$time_dif_z #Using the velocity formula

#Plotting z velocity against time
plot(data$time, data$velocity_z)

ggplot(data, aes(time, z))+
  geom_path()+ #adding geom to track path
  theme_bw()+ #Changing background colour
  labs(y="Velocity on z-axis", x = "Time") #Adding labels

```

#Apply a Butterworth filter to z velocity with reasonable parameters. Plot the filtered velocity on top of the unfiltered one
```{r applying butterworth filter}
bf <- butter(4, 0.1, type ='low') #defining butterworth filter with values used in Vesper et al. (2017)

data$velocity_z_f <- filtfilt(bf,data$velocity_z) #Applying filter to z variable


#Plotting filtered velocity on top of unfiltered
ggplot(data, aes(time, velocity_z_f))+ 
  geom_path(aes(col = "Filtered"))+ #adding geom to track path of filtered z velocity
  geom_path(aes(time, velocity_z, col = "Unfiltered")) + #adding geom to track path of unfilteres z velocity
  theme_bw()+ #Changing background colour
  labs(y="Velocity on z-axis", x = "Time") + #Adding labels
  scale_colour_manual(name="Velocity",
    values=c(Filtered="red", Unfiltered="blue")) +  #Adding legend
  ggtitle("Velocity of Z") #Adding title
  


```


#Make a final plot with z and the nicely filtered velocity over time.
```{r}
#Plotting filtered velocity on top of unfiltered
ggplot()+
  geom_path(data = data, aes(time, velocity_z_f, col = "Velocity"))+ #adding geom to trace smoothed velocity over time
  geom_path(data = data, aes(time, flipped_z, col = "Position")) + #adding geom to track z-position over time
  theme_bw()+ #Changing background colour
  labs(y="Velocity and position, respectively", x = "Time") + #Adding labels 
  scale_colour_manual(name="Colour",
    values=c(Position="red", Velocity="blue")) +  #Adding legend and colours
  ggtitle("Velocity and Z-position over time") #Adding title

```



